---
title: "Footprint Calculation"
author: "Cebulski, A.C"
date: "`r Sys.Date()`"
output: html_document
bibliography: [../../papers/bibtex/library.bib]
csl: ../../papers/bibtex/apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(weatherdash)
```

TODO:

- lower EC was raised winter 2022 need to fix in QAQC before applying across all dates
- instrument height changes relative to snow calculate using SR50 data? 
- what is the difference in  "horiz_wind_spd" & "result_wind_spd"

## Formulas

Calculate the upwind area aka footprint ($X_f$, m) representative of these measurements that accounts for a percentage of the flux. Where $U$ is the assumed constant wind speed, $u^{*}$ is the friction velocity, $d$ is the displacement height due to vegetation (taken as 2/3 the vegetation height), $z$ is the instrument height, $k$ is the von karmans constant (usually taken as 0.4), $f$ is the fraction of the area that explains the flux [@Schuepp1990]:

$$
X_f = - \frac{1}{ln(f)}\frac{U}{u^*}\frac{z-d}{k}
$$ 
Friction velocity ($u_*$) can be modelled as in wind_profile.Rmd or is also measured at the Eddy Covariance Sites. 

## find the parameters $u^{∗}$, $Z_{0m}$, and $d_0$ of the logarithmic wind profile equation that best matches the measured data provided below:

Constants:

```{r}
# need to measure this in the field still (2022-06-01)
temp_offset <- 125

k <- 0.4 # von karmans constant (unitless) 

d0 <- 1.965869 # m estimated from jan 16 2022 at 4:00 pm 
f <- 0.8 # 80%

```

Load Data: 

```{r}
low_wind <- readRDS('data/eddy_cov_15min_combined.rds') |> select(TIMESTAMP, wnd_dir_compass, u_star, low_ec_rslt_wnd_spd = result_wind_spd)# |> 
#   pivot_longer(u_star:low_ec_rslt_wnd_spd)
# 
# ggplot(low_wind, aes(TIMESTAMP, value, colour = name)) + geom_line()
# 
# plotly::ggplotly()

mid_hi_wind <- readRDS('../../analysis/treefort-main/data/met/met_main.rds') |> select(TIMESTAMP, USWindSpeed_S_WVT, top_ec_rlst_wnd_spd, top_ec_u_star)

wind_wide <- left_join(low_wind, mid_hi_wind, by = 'TIMESTAMP')

wind_dir <- wind_wide |> 
  mutate(wind_dir = (wnd_dir_compass + temp_offset) %% 365) |> 
  select(TIMESTAMP, wind_dir) 

low <- wind_wide |> 
  select(TIMESTAMP, u_star, wind_speed = low_ec_rslt_wnd_spd) |> 
  mutate(inst_height = 3) |> 
  mutate(
    med = zoo::rollapply(wind_speed, width = 1000, FUN = median, na.rm = T, align = 'center', partial = T),
    sd = zoo::rollapply(wind_speed, width = 1000, FUN = sd, na.rm = T, align = 'center', partial = T),
    stdep = (wind_speed - med) / sd,
    flag = ifelse(wind_speed > 4, T, F)
  ) |> 
    select(TIMESTAMP, u_star, wind_speed, inst_height, flag) 


ggplot(low, aes(TIMESTAMP, wind_speed, colour = flag)) + geom_point()

plotly::ggplotly()

hi <- wind_wide |> 
  select(TIMESTAMP, u_star = top_ec_u_star, wind_speed = top_ec_rlst_wnd_spd) |> 
  mutate(
    inst_height = 13.5,
    mean = zoo::rollapply(wind_speed, width = 48, FUN = mean, na.rm = T, align = 'center', partial = T),
    sd = zoo::rollapply(wind_speed, width = 48, FUN = sd, na.rm = T, align = 'center', partial = T),
    stdep = (wind_speed - mean) / sd,
    flag = ifelse(wind_speed > 20, T, F)
  ) |> 
  select(TIMESTAMP, u_star, wind_speed, inst_height, flag) 

ggplot(hi, aes(TIMESTAMP, wind_speed, colour = flag)) + geom_point()
plotly::ggplotly()


all <- rbind(low, hi) |> 
  left_join(wind_dir) |> 
  filter(flag == F)

all <- weatherdash::cat_wind(all, 'wind_dir', 'wind_dir_bin')

ggplot(all, aes(TIMESTAMP, wind_speed, colour = flag)) + 
  geom_point() +
  facet_wrap(~inst_height)
plotly::ggplotly()
  
#|> filter(as.Date(TIMESTAMP) == as.Date('2022-01-16')) 

```

Calculate $X_f$:

```{r}
x_f <- all |> 
  mutate(x_f = -(1/log(f))*(wind_speed/u_star)*((inst_height-d0)/k)) |> 
  filter(x_f < 5000)

```

Plot $X_f$ for 3m EC with wind speed on all days: 
NOTE EC WAS RAISED Feb 17, 2022, this could be why relationship changes after FEB plot.

```{r}
ggplot(x_f |> filter(inst_height == 3), aes(wind_speed, x_f, colour = wind_dir_bin)) +
  geom_point() +
  facet_wrap(~inst_height+strftime(as.Date(TIMESTAMP), "%m"))
```

Plot $X_f$ for 13.5 m EC with wind speed on all days: 

Maybe larger footprint during winter months when snow is in the trees? 

```{r}
ggplot(x_f |> filter(inst_height == 13.5), aes(wind_speed, x_f, colour = wind_dir_bin)) +
  geom_point() +
  facet_wrap(~inst_height+strftime(as.Date(TIMESTAMP), "%m"), scales = 'free')

plotly::ggplotly()

```

Output $X_f$ table

```{r}
x_f |> 
  mutate(yr_mon =  strftime(as.Date(TIMESTAMP), "%y-%m")) |> 
  group_by(inst_height, yr_mon) |> 
  summarise(x_f_mean = mean(x_f, na.rm = T),
            u_star_mean = mean(u_star, na.rm = T),
            wind_spd_mean = mean(wind_speed, na.rm = T),
            wind_spd_pk_mean = max(wind_speed, na.rm = T),
            wind_dir_mean = mean(wind_dir, na.rm = T))
```


Wind Rose for all months for each instrument height:

```{r}
# for (i in c(3,13.5)) {
#   for (m in 1:12) {
#     fltr <- all |> 
#       mutate(month = as.numeric(strftime(as.Date(TIMESTAMP), "%m"))) |> 
#       filter(inst_height == i,
#              month == m)
#     
#   if(nrow(fltr) != 0){
#     p <- weatherdash::wind_rose(fltr, 'TIMESTAMP', 'wind_speed', 'wind_dir', plot_title = paste('Height', i, "m", 'for', m))
#     file1 <- paste0('figs/wind_rose/wind_rose_height_', i, '_', m,'.html')
#     file2 <- paste0('figs/wind_rose/wind_rose_height_', i, '_', m,'.png')
# 
#     htmlwidgets::saveWidget(plotly::as_widget(p), file1)
#     plotly::save_image(p, file2)
#   }
# 
#   }
#   
# }
```

Look at one of the windiest days without snow on instruments:

```{r}
ggplot(x_f |> filter(as.Date(TIMESTAMP) == as.Date('2022-01-16')) , aes(wind_speed, x_f)) +
  geom_point() +
  facet_wrap(~inst_height, scales = 'free')
```

Wind Rose on this Day:

```{r}


low_wind <- readRDS('data/eddy_cov_15min_combined.rds') |> 
  filter(as.Date(TIMESTAMP) == as.Date('2022-01-16')) |> 
  select(TIMESTAMP, 
         wnd_dir_compass, 
         u_star, 
         low_ec_rslt_wnd_spd = result_wind_spd) |>
  mutate(
    wind_dir = (wnd_dir_compass + temp_offset) %% 365,
    mean = zoo::rollapply(low_ec_rslt_wnd_spd, width = 48, FUN = mean, na.rm = T, align = 'center', partial = T),
    sd = zoo::rollapply(low_ec_rslt_wnd_spd, width = 48, FUN = sd, na.rm = T, align = 'center', partial = T),
    stdep = (low_ec_rslt_wnd_spd - mean) / sd,
    flag = ifelse(low_ec_rslt_wnd_spd > 4, T, F)
  ) 

ggplot(low_wind, aes(TIMESTAMP, low_ec_rslt_wnd_spd, colour = flag)) + 
  geom_point() 

weatherdash::wind_rose(low_wind, 'TIMESTAMP', 'low_ec_rslt_wnd_spd', 'wind_dir', plot_title = 'test')
```

